Biogeography and Environmental Conditions Shape Phage and Bacteria Interaction Networks Across the Healthy Human Microbiome

### 2017-03-28 doc/manuscript.Rmd

We reported that, just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time. Gut network structure was not conserved among family members. These findings suggest that microbiome properties such as stability and the potential for horizontal gene transfer are personal, and potentially impacted by personal factors ranging from the body's immune system to environemntal conditions such as climate and diet. The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were assocaited with altered network structures.

@@

## Centrality Analysis
Degree and closeness centrality were calculated using the associated functions within the igraph R package (v1.0.1) [@Theigraphsoftware:vh]. Briefly, the **closeness centrality** of Vi was calculated taking the inverse of the average length of the shortest path (`d`) between nodes Vi and the `k` other nodes Vj, such that closeness centrality of node Vi is:

$$ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ k }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } $$

The distance between nodes (`d`) was calculated as the shortest number of edges required to be traversed to move from one node to another.

The simple metric of **degree centrality** of node Vi was defined as the sum of the number of k edges Ei associated with that node, such that:

$$ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ i=1 }^{ k }{ { E }_{ i } } $$

The Eigenvector centrality was calculated using the values of the first eigen vector of the associated adjacency matrix that are associated with each vertex `Vi`. Conceptually, this function results in a centrality value that reflects the connections of the vertex, as well as the centrality of the connected vertices.

The **centralization** metric was used to assess the average centrality of each sample graph `G`. Centralization was calculated by taking the sum of each vertex `Vi` centrality from the graph maximum centrality `Cw`, such that:

$$ C\left( G \right) =\frac { \sum _{ i=1 }^{ k }{ \max _{ w }{ c\left( w \right)  } -c\left( { v }_{ i } \right)  }  }{ { T } }  $$

@@

@@

Where `Cij` is the sum of the lesser centrality values for those vertices shared between graphs, and `Ci` and `Cj` are are the total number of vertices found in each graph. This allows us calculate the dissimilarity between graphs based on the shared centrality values between the two graphs.

>>>>

We reported that, just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time. Gut network structure was not conserved among family members. These findings suggest that microbiome properties such as stability and the potential for horizontal gene transfer are personal, and potentially impacted by personal factors ranging from the body's immune system to environemntal conditions such as climate and diet. The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.

@@

## Centrality Analysis
Let $G(V,E)$ be an undirected, unweighted graph with $|V|=n$ nodes and $|E|=m$ edges. Also, let $\mathbf{A}$ be its corresponding adjacency matrix with entries $a_{ij} = 1$ if nodes $V_i$ and $V_j$ are connected via an edge, and $a_{ij} = 0$ otherwise.

Briefly, the **closeness centrality** of node $V_i$ is calculated taking the inverse of the average length of the shortest paths (`d`) between nodes $V_i$ and all the other nodes $V_j$. Mathematically, the closeness centrality of node $V_i$ is given as:

$$ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ n }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } $$

[//]: <> The distance between nodes (`d`) was calculated as the shortest number of edges required to be traversed to move from one node to another.

Intuitively, the **degree centrality** of node $V_i$ is defined as the number of edges that are adjacent to that node:

$$ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ j=1 }^{ n }{ { a }_{ ij } } $$

where $a_{ij}$ is the $ij^{th}$ entry in the adjacency matrix $\mathbf{A}$.

[//]: <> The simple metric of degree centrality of node Vi was defined as the sum of the number of k edges Ei associated with that node, such that:

The Eigenvector centrality of node $V_i$ is defined as the $i^{th}$ value in the first eigenvector of the associated adjacency matrix $\mathbf{A}$. Conceptually, this function results in a centrality value that reflects the connections of the vertex, as well as the centrality of its neighboring vertices.

The **centralization** metric was used to assess the average centrality of each sample graph `G`. Centralization was calculated by taking the sum of each vertex `Vi`'s centrality from the graph maximum centrality `Cw`, such that:

$$ C\left( G \right) =\frac { \sum _{ i=1 }^{ n }{ Cw -c\left( { v }_{ i } \right)  }  }{ { T } }  $$

@@

Degree and closeness centrality were calculated using the associated functions within the igraph R package (v1.0.1) [@Theigraphsoftware:vh].

@@

Where `Cij` is the sum of the lesser centrality values for those vertices shared between graphs, and `Ci` and `Cj` are the total number of vertices found in each graph. This allows us to calculate the dissimilarity between graphs based on the shared centrality values between the two graphs.
